---
title: "Socialeconomic Status and Weight-related Perception & Behavior"
author: "Songyi Ahn"
format: 
  revealjs:
    embeded-resources: true
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, eval=TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(rio)
library(here)
library(tidyverse)
library(ggridges)
library(ggmosaic)
```

## Background

-   Topic
-   Data source and content

------------------------------------------------------------------------

## Research Questions

-   RQ1: How is socioeconomic status (SES) associated with an individual’s **perceptions of their own weight**?
-   RQ2: How is socioeconomic status (SES) associated with the frequency of **calorie label checking for weight** management?
-   RQ3: How is socioeconomic status (SES) associated with engagement in various **weight loss behaviors**?
-   RQ4: How is socioeconomic status (SES) associated with individuals’ choices of **dining locations**?

------------------------------------------------------------------------

## Brief Showcase of Visualizations

::: incremental
### SES and Weight Perception

```{r}
weight1 <- import(here("data", "sesdata.rds"))
```

```{r}
weight1 %>%
ggplot(aes(x = SES, y = BMI, color = weight_perception)) +
  geom_jitter(width = 0.2, alpha = 0.7) +
  ylim(NA, 50) +
  scale_color_viridis_d() +
  theme_minimal() +
  scale_x_continuous(breaks = c(1, 4),
                     labels = c("Low", "High")) +
  labs(x = "SES", y = "BMI", color = "Weight Perception",
       title = "Differences in Weight Perception by SES at the Same BMI")
```
:::

::: incremental
### SES and Calorie Label Checking for Weight

```{r}
weight2 <- import(here("data", "sesdata.rds"))
```

```{r}
weight2 %>%
  mutate(
    SES_group = cut(SES, 
                  breaks = quantile(SES, probs = c(0, 1/3, 2/3, 1), na.rm = TRUE),
                  include.lowest = TRUE,
                  labels = c("Low", "Middle", "High")),
    calorie_label_usage = recode(calorie_label_usage,
                                 `1` = "Always",
                                 `2` = "Most of the time",
                                 `3` = "Sometimes",
                                 `4` = "Rarely",
                                 `5` = "Never"),
    calorie_label_usage = factor(calorie_label_usage, 
                                 levels = c("Never", "Rarely", "Sometimes", "Most of the time", "Always"))
  ) %>% 
  count(SES_group, calorie_label_usage) %>%
  ggplot(aes(x = SES_group, y = calorie_label_usage, fill = n)) + 
  geom_tile(color = "white") + 
  scale_fill_viridis_c(option = "D") +
  theme_minimal() +
  labs(
    x = "SES",
    y = "Frequency of Calorie Label Checking for Weight Management",
    fill = "Count",
    title = "Calorie Label Checking by SES Group"
  )
```
:::

::: incremental
### SES and Weight Loss Behaviors

```{r}
weight3 <- import(here("data", "sesdata.rds"))
```

```{r}
weight3 %>%
  ggplot(aes(x = SES, y = weight_loss_behavior_type, fill = weight_loss_behavior_type)) +
  geom_density_ridges(alpha = 0.8) +  
  scale_fill_viridis_d() +
  theme_minimal() +
  stat_summary(
    aes(label = weight_loss_behavior_type),  
    fun = median, 
    geom = "text",
    color = "white",  
    size = 2,  
    hjust = 0.8,  
    vjust = -3
  ) +
  theme(axis.text.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  scale_x_continuous(breaks = c(1, 4),
                     labels = c("Low", "High")) +
  labs(x = "SES", y = "Weight Loss Behavior Type", fill = "Weight Loss Method",
       title = "Variation in Weight Loss Methods Across SES Groups")
```
:::

::: incremental
### SES and Dining Location Choice

```{r}
weight4 <- import(here("data", "sesdata.rds"))
```

```{r}
weight4$SES_group2 <- cut(weight4$SES,
                         breaks = quantile(weight4$SES, probs = c(0, 0.5, 1), na.rm = TRUE),
                         include.lowest = TRUE,
                         labels = c("Low", "High"))

weight_summary4 <- weight4 %>%
  count(SES_group2, dininglocation)

weight_summary4$dininglocation <- factor(weight_summary4$dininglocation, 
                                         levels = c("fastfood", "buffet", "restaurant"))

ggplot(weight_summary4) +
  geom_mosaic(aes(weight = n, x = product(SES_group2), fill = dininglocation)) +
  scale_fill_viridis_d() +
  theme_minimal() +
  labs(x = "SES", y = "Proportion of Dining Location", fill = "Dining Location",
       title = "Dining Preferences by Socioeconomic Status",
       caption = "Respondents could select multiple dining locations")
```
:::

## In-Depth Discussion: Plot 1. SES and Weight Perception

::: incremental
### Intended Audience

-   Intended audience for this plot is blah blah blah 
:::

::: incremental
### Design Choice

```{r}
weight1 %>%
ggplot(aes(x = SES, y = BMI, color = weight_perception)) +
  geom_jitter(width = 0.2, alpha = 0.7) +
  ylim(NA, 50) +
  scale_color_viridis_d() +
  theme_minimal() +
  scale_x_continuous(breaks = c(1, 4),
                     labels = c("Low", "High")) +
  labs(x = "SES", y = "BMI", color = "Weight Perception",
       title = "Differences in Weight Perception by SES at the Same BMI")
```

-   Type of plot
-   Layout
-   Colors
:::

::: incremental
### Prior version & Change
```{r}
weight1 %>%
ggplot(aes(x = SES, y = BMI, color = weight_perception)) +
  geom_jitter(width = 0.2, alpha = 0.7) +
  theme_minimal()
```

-   y limit
-   color change
-   title, labs 
:::

::: incremental
### Encountered Challenge

-   Can't really interpret 
:::

::: incremental
### Victory

-   I anyhow made it
:::

## In-Depth Discussion: Plot 2. SES and Calorie Label Checking for Weight

::: incremental
### Intended Audience

-   Intended audience for this plot is blah blah blah 
:::

::: incremental
### Design Choice

```{r}
weight2 %>%
  mutate(
    SES_group = cut(SES, 
                  breaks = quantile(SES, probs = c(0, 1/3, 2/3, 1), na.rm = TRUE),
                  include.lowest = TRUE,
                  labels = c("Low", "Middle", "High")),
    calorie_label_usage = recode(calorie_label_usage,
                                 `1` = "Always",
                                 `2` = "Most of the time",
                                 `3` = "Sometimes",
                                 `4` = "Rarely",
                                 `5` = "Never"),
    calorie_label_usage = factor(calorie_label_usage, 
                                 levels = c("Never", "Rarely", "Sometimes", "Most of the time", "Always"))
  ) %>% 
  count(SES_group, calorie_label_usage) %>%
  ggplot(aes(x = SES_group, y = calorie_label_usage, fill = n)) + 
  geom_tile(color = "white") + 
  scale_fill_viridis_c(option = "D") +
  theme_minimal() +
  labs(
    x = "SES",
    y = "Frequency of Calorie Label Checking for Weight Management",
    fill = "Count",
    title = "Calorie Label Checking by SES Group"
  )
```

-   Type of plot
-   Layout
-   Colors
:::

::: incremental
### Prior version & Change
```{r}
weight2 %>%
  mutate(SES_bin = cut(SES, breaks = 3)) %>% 
  ggplot(aes(x = SES_bin, y = factor(calorie_label_usage))) +
  geom_bin2d() +
  scale_fill_gradient(low = "white", high = "blue") +
  theme_minimal() 
```

-   SES group divide by quantile 
-   color change
-   Geom_bin2d to geom_tile
:::

::: incremental
### Encountered Challenge

-   color not instinctive to in grasp the meaning?
:::

::: incremental
### Victory

-   I anyhow made it
-   renaming frequency number and factor ordering 
:::