---
title: "Visualizations for Peer Review"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, eval=TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(rio)
library(here)
library(tidyverse)
library(ggmosaic)
```

```{r}
weight <- import(here("data", "cleaned_data.rds"))
```

# income, weight perception, BMI
```{r}
ggplot(weight, aes(x = income_level, y = BMI, color = weight_perception)) +
  geom_jitter(alpha = 0.5, width = 0.2) +
  ylim(NA, 40) +
  theme_minimal()
```

# education level, weight perception, BMI
```{r}
ggplot(weight, aes(x = education_level, y = BMI, color = weight_perception)) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  ylim(NA, 40) +
  theme_minimal() +
  labs(x = "Education Level", y = "BMI", color = "Weight Perception")
```

# health insurance type, weight perception, BMI
```{r}
weight$health_insurance_type <- factor(weight$health_insurance_type, 
                                        levels = c("state_sponsored_plan", "medicaid", "private_insurance"))

ggplot(weight, aes(x = health_insurance_type, y = BMI, color = weight_perception)) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  ylim(NA, 40) +
  theme_minimal() +
  labs(x = "Types of Health Insurance", y = "BMI", color = "Weight Perception")
```


# income and dining choice 
```{r}
weight$income_group <- cut(weight$income_level, 
                           breaks = c(0, 2.5, 5), 
                           labels = c("Low Income", "High Income"),
                           include.lowest = TRUE)

weight_summary <- weight %>%
  count(income_group, education_level, health_insurance_type, dininglocation)

weight_summary$dininglocation <- factor(weight_summary$dininglocation, 
                                        levels = c("fastfood", "buffet", "restaurant"))

ggplot(weight_summary) +
  geom_mosaic(aes(weight = n, x = product(income_group), fill = dininglocation)) +
  scale_fill_manual(values = c("fastfood" = "lightgreen", 
                               "buffet" = "green", 
                               "restaurant" = "darkgreen")) +
  theme_minimal()
```

# education level and dining choice
```{r}
ggplot(weight_summary) +
  geom_mosaic(aes(weight = n, x = product(education_level), fill = dininglocation)) +
  scale_fill_manual(values = c("fastfood" = "lightgreen", 
                               "buffet" = "green", 
                               "restaurant" = "darkgreen")) +
  theme_minimal()
```

# health insurance type and dining choice
```{r}
weight_summary$health_insurance_type <- factor(weight_summary$health_insurance_type, 
                                               levels = c("state_sponsored_plan", "medicaid", "private_insurance"))

ggplot(weight_summary) +
  geom_mosaic(aes(weight = n, x = product(health_insurance_type), fill = dininglocation)) +
  scale_fill_manual(values = c("fastfood" = "lightgreen", 
                               "buffet" = "green", 
                               "restaurant" = "darkgreen")) +
  theme_minimal()
```

# income and label calories check for weight control
```{r}
ggplot(weight, aes(x = income_level, y = calorie_label_usage)) +
  geom_jitter(alpha = 0.2) +  
  geom_smooth(method = "lm", color = "red") +
  theme_minimal() +
  labs(x = "Income Level", y = "Calorie Label Checking Frequency for Weight Control")
```

# education level and label calories check for weight control
```{r}
ggplot(weight, aes(x = education_level, y = calorie_label_usage)) +
  geom_jitter(alpha = 0.2) +  
  geom_smooth(method = "lm", color = "blue") +
  theme_minimal() +
  labs(x = "Education Level", y = "Calorie Label Checking Frequency for Weight Control")

```

# health insurance type and label calories check for weight control
```{r}
ggplot(weight, aes(x = health_insurance_type, y = calorie_label_usage)) +
  geom_jitter(alpha = 0.2) +  
  theme_minimal() +
  labs(x = "Health Insurance Type", y = "Calorie Label Checking Frequency for Weight Control")
```

# income and weight loss behavior
```{r}
ggplot(weight, aes(x = income_level, fill = weight_loss_behavior_type, color = weight_loss_behavior_type)) +
  geom_density(alpha = 0.5) +  
  scale_fill_viridis_d() + 
  theme_minimal() +
  labs(x = "Income Level", y = "Density", fill = "Weight Loss Behavior Type", color = "Weight Loss Behavior Type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

# education level and weight loss behavior
```{r}
ggplot(weight, aes(x = education_level, fill = weight_loss_behavior_type)) +
  geom_bar(position = "fill") +  
  scale_fill_viridis_d() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# health insurance type and weight loss behavior
```{r}
ggplot(weight, aes(x = health_insurance_type, fill = weight_loss_behavior_type)) +
  geom_bar(position = "fill") +  
  scale_fill_viridis_d() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

